---
title: "Hypo_tests_midterm"
author: "Sahara Ensley"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```
AdoptionSpeed - Categorical speed of adoption. Lower is faster. This is the value to predict. See below section for more info.
Type - Type of animal (1 = Dog, 2 = Cat)
Age - Age of pet when listed, in months
Gender - Gender of pet (1 = Male, 2 = Female, 3 = Mixed, if profile represents group of pets)
MaturitySize - Size at maturity (1 = Small, 2 = Medium, 3 = Large, 4 = Extra Large, 0 = Not Specified)
FurLength - Fur length (1 = Short, 2 = Medium, 3 = Long, 0 = Not Specified)
Vaccinated - Pet has been vaccinated (1 = Yes, 2 = No, 3 = Not Sure)
PhotoAmt - Total uploaded photos for this pet
## Loading data
```{r}
# dat = read.csv('/Users/sahara/Documents/GW/DataScience/petfinder-adoption-prediction/train/train.csv')
data = read.csv('/Users/sahara/Documents/GW/DataScience/data_torture/chapter_5/datafile.csv')
data$AdoptionSpeed <- factor(data$AdoptionSpeed, order=T, levels = c(0,1,2,3,4))

# converting the data types we want
data$Type = as.factor(data$Type)
data$MaturitySize = as.factor(data$MaturitySize)
data$FurLength = as.factor(data$FurLength)
data$Vaccinated = as.factor(data$Vaccinated)
data$Gender = as.factor(data$Gender)

# Only looking at profiles with 1 animal
data = subset(data, data$Quantity == 1)
data = subset(data, ! data$AdoptionSpeed == 4)

# Only pulling the columns we want to look at
dat = data[c('Type', 'Age', 'Gender', 'MaturitySize', 'FurLength', 'Vaccinated', 'PhotoAmt', 'AdoptionSpeed')]
summary(dat)
```

```{r}
# Adding random values
dat$ASnum = NaN
dat$ASnum[dat$AdoptionSpeed == 0] = 0
dat$ASnum[dat$AdoptionSpeed == 1] = runif(sum(dat$AdoptionSpeed == 1), 1, 7)
dat$ASnum[dat$AdoptionSpeed == 2] = runif(sum(dat$AdoptionSpeed == 2), 8, 30)
dat$ASnum[dat$AdoptionSpeed == 3] = runif(sum(dat$AdoptionSpeed == 3), 31, 90)
#dat$ASnum[dat$AdoptionSpeed == 4] = 100

summary(dat)
```


```{r}
hist(data = dat, x = dat$ASnum)
```

```{r}
qqnorm(y = dat$ASnum)
```


## SMART: Do dogs get adopted faster than cats?
```{r}
library('ggplot2')
#pdf('type_boxplot.pdf')
ggplot(data = dat, aes(y = ASnum, x = Type, fill = Type)) + 
  geom_boxplot() +
  labs(title = "Adoption Speed by Animal Type", x = 'Animal Type', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Dog", "2" = "Cat")) + 
  scale_fill_discrete(labels = c("Dog", "Cat"))

#dev.off()
```

```{r}
bp_dogscats = bptest(data = dat, formula = ASnum ~ Type)
bp_dogscats
dogs = subset(dat, dat$Type == 1)
cats = subset(dat, dat$Type == 2)

dc_t = t.test(dogs$ASnum, cats$ASnum)
dc_t
```



## SMART: Do physical attributes affect adoption speed?
```{r}
### GENDER
#pdf('Gender_boxplot.pdf')
bp_mf = bptest(data = dat, formula = ASnum ~ Gender)
bp_mf
male = subset(dat, dat$Gender == 1)
female = subset(dat, dat$Gender == 2)
mf_t = t.test(male$ASnum, female$ASnum)
mf_t


ggplot(data = dat, aes(y = ASnum, x = Gender, fill = Gender)) + 
  geom_boxplot() + 
  labs(title = "Adoption Speed by Animal Gender", x = 'Animal Gender', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Male", "2" = "Female")) + 
  scale_fill_discrete(labels = c("Male", "Female"))
#dev.off()
```

```{r}
### SIZE
bp_size = bptest(data = dat, formula = ASnum ~ MaturitySize)
bp_size
#pdf('Size_boxplot.pdf')
ggplot(data = dat, aes(y = ASnum, x = MaturitySize, fill = MaturitySize)) + 
  geom_boxplot() + 
  labs(title = "Adoption Speed by Maturity Size", x = 'Maturity Size', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Small", "2" = "Medium", "3" = "Large", "4" = "Extra Large")) + 
  scale_fill_discrete(labels = c("Small", "Medium", "Large", "Extra Large"))
#dev.off()
anova_size = anova(aov(data = dat, ASnum ~ MaturitySize))
anova_size
turkey_size = TukeyHSD(aov(data = dat, ASnum ~ MaturitySize))
turkey_size
```

```{r}
### FURLENGTH
bp_fur = bptest(data = dat, formula = ASnum ~ FurLength)
bp_fur

#pdf('fur_boxplot.pdf')
ggplot(data = dat, aes(y = ASnum, x = FurLength, fill = FurLength)) + 
  geom_boxplot() +
  labs(title = "Adoption Speed by Fur Length", x = 'Fur Length', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Short", "2" = "Medium", "3" = "Long")) + 
  scale_fill_discrete(labels = c("Short", "Medium", "Long"))
#dev.off()
anova_fur = anova(aov(data = dat, ASnum ~ FurLength))
anova_fur
turkey_fur = TukeyHSD(aov(data = dat, ASnum ~ FurLength))
turkey_fur
```


```{r}
### VACCINATED
bp_vac = bptest(data = dat, formula = ASnum ~ Vaccinated)
bp_vac
#pdf('vacc_boxplot.pdf')
ggplot(data = dat, aes(y = ASnum, x = Vaccinated, fill = Vaccinated)) + 
  geom_boxplot() + 
  labs(title = "Adoption Speed by Vaccination Status", x = 'Vaccination Status', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Vaccinated", "2" = "Unvaccinated", "3" = "Not Specified")) + 
  scale_fill_discrete(labels = c("Vaccinated", "Unvaccinated", "Not Specified"))
#dev.off()
anova_vac = anova(aov(data = dat, ASnum ~ Vaccinated))
anova_vac
turkey_vac = TukeyHSD(aov(data = dat, ASnum ~ Vaccinated))
turkey_vac
```


## SMART: Do profile attributes affect adoption speed?

```{r}
### PHOTOS
bp_pho = bptest(data = dat, formula = ASnum ~ PhotoAmt)
bp_pho

ggplot(data = dat, aes(y = ASnum, fill = PhotoAmt)) + 
  geom_boxplot()

anova_pho = anova(aov(data = dat, ASnum ~ PhotoAmt))
anova_pho
turkey_pho = TukeyHSD(aov(data = dat, ASnum ~ PhotoAmt))
turkey_pho
```

## SMART: Do more expensive dogs get adopted slower?
```{r}

histogram(data = dogs, x = dogs$Fee)
qqnorm(y = dogs$Fee)
ggplot(data = dogs, aes(x = Fee, y = ASnum)) + 
  geom_point() + 
  geom_smooth(method = lm)
```

```{r}
an2 = anova(aov(data = dogs, ASnum ~ Fee))
an2
```


ok let's remove that outlier now

```{r}
dogs_noout = outlierKD2(dogs, var = Fee, rm = TRUE)


histogram(data = dogs_noout, x = dogs_noout$Fee)
qqnorm(y = dogs_noout$Fee)
ggplot(data = dogs_noout, aes(x = Fee, y = ASnum)) + 
  geom_point() + 
  geom_smooth(method = lm)
```

```{r}
an2_noout = anova(aov(data = dogs_noout, ASnum ~ Fee))
an2_noout
```

okay so everything that wasn't 0 was deemed an outlier ...

```{r}
## ORDINAL
contable = table(dat$AdoptionSpeed, dat$Age)
chitest = chisq.test(contable)
chitest
contable
```

## SMART: Do dogs with more pictures get adopted faster?

```{r}
histogram(data = dogs, x = dogs$PhotoAmt)
qqnorm(dogs$PhotoAmt)
```


```{r}
an3 = anova(aov(data = dogs, ASnum ~ PhotoAmt))
an3
```

## SMART: Do dogs with more videos get adopted faster?

```{r}
histogram(data = dogs, x = dogs$VideoAmt)
qqnorm(dogs$VideoAmt)
```


```{r}
an3 = anova(aov(data = dogs, ASnum ~ VideoAmt))
an3
```


