---
title: "EDA_pawpularity"
author: "Alexis"
date: "whenever we finish"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
library(ezids)
library(ggplot2)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
options(ggplot2.continuous.color = "viridis")
options(ggplot2.continuous.fill = "viridis")
options(ggplot2.discrete.fill = "viridis")
```


```{r results=T}
pawp <- data.frame(read.csv("pawpularity.csv"))
str(pawp)
```
## Explaining different variables
- Subject.Focus: Pet stands out against uncluttered background, not too close / far.
- Eyes: Both eyes are facing front or near-front, with at least 1 eye / pupil decently clear.
- Face: Decently clear face, facing front or near-front.
- Near: Single pet taking up significant portion of photo (roughly over 50% of photo width or height).
- Action: Pet in the middle of an action (e.g., jumping).
- Accessory: Accompanying physical or digital accessory / prop (i.e. toy, digital sticker), excluding collar and leash.
- Group: More than 1 pet in the photo.
- Collage: Digitally-retouched photo (i.e. with digital photo frame, combination of multiple photos).
- Human: Human in the photo.
- Occlusion: Specific undesirable objects blocking part of the pet (i.e. human, cage or fence). Note that not all blocking objects are considered occlusion.
- Info: Custom-added text or labels (i.e. pet name, description).
- Blur: Noticeably out of focus or noisy, especially for the pet’s eyes and face. For Blur entries, “Eyes” column is always set to 0.
- Pawpularity: The Pawpularity Score is derived from each pet profile's page view statistics at the listing pages, using an algorithm that normalizes the traffic data across different pages, platforms (web & mobile) and various metrics. Duplicate clicks, crawler bot accesses and sponsored profiles are excluded from the analysis.
