---
title: "Hypo_tests_midterm"
author: "Sahara Ensley"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```
AdoptionSpeed - Categorical speed of adoption. Lower is faster. This is the value to predict. See below section for more info.\
Type - Type of animal (1 = Dog, 2 = Cat)\
Age - Age of pet when listed, in months\
Gender - Gender of pet (1 = Male, 2 = Female, 3 = Mixed, if profile represents group of pets)\
MaturitySize - Size at maturity (1 = Small, 2 = Medium, 3 = Large, 4 = Extra Large, 0 = Not Specified)\
FurLength - Fur length (1 = Short, 2 = Medium, 3 = Long, 0 = Not Specified)\
Vaccinated - Pet has been vaccinated (1 = Yes, 2 = No, 3 = Not Sure)\
PhotoAmt - Total uploaded photos for this pet\
## Intro
```{r}
library("corrplot")
library("ezids")
library("gtsummary")
## Loading data
data = read.csv('chapter_5/datafile.csv')
data$AdoptionSpeed <- factor(data$AdoptionSpeed, order=T, levels = c(0,1,2,3,4))

# converting the data types we want - categorical
data$Type = as.factor(data$Type)
data$MaturitySize = as.factor(data$MaturitySize)
data$FurLength = as.factor(data$FurLength)
data$Vaccinated = as.factor(data$Vaccinated)
data$Gender = as.factor(data$Gender)

# Only looking at profiles with 1 animal
data = subset(data, data$Quantity == 1)
data = subset(data, ! data$AdoptionSpeed == 4)

# Only pulling the columns we want to look at
dat = data[c('Type', 'Age', 'Gender', 'MaturitySize', 'FurLength', 'Vaccinated', 'PhotoAmt','VideoAmt', 'AdoptionSpeed')]
summary(dat)

# Adding random values
dat$ASnum = NaN
dat$ASnum[dat$AdoptionSpeed == 0] = 0
dat$ASnum[dat$AdoptionSpeed == 1] = floor(runif(sum(dat$AdoptionSpeed == 1), 1, 8))
dat$ASnum[dat$AdoptionSpeed == 2] = floor(runif(sum(dat$AdoptionSpeed == 2), 8, 31))
dat$ASnum[dat$AdoptionSpeed == 3] = floor(runif(sum(dat$AdoptionSpeed == 3), 31, 91))
#dat$ASnum[dat$AdoptionSpeed == 4] = 100 -- this line isn't necessary because we exclude animals that have this adoption speed

summary(dat)
```

## Chapter 2
```{r}
str(dat)
```


## Chapter 3
### Numerical variable
In this chapter we will introduce some variables. The first one is AdoptionSpeed.  
```{r histogram of AdoptionSpeed}
loadPkg("ggplot2")
#Histogram
ggplot(data=dat, aes(AdoptionSpeed)) + 
  geom_bar(breaks=seq(0, 4, by=1),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="AdoptionSpeed of pets", x = "AdoptionSpeed") + 
  scale_fill_gradient("Count", low="green", high="red")

```
AdoptionSpeed is divided into 4 levels.   
0 means pet was adopted on the same day as it was listed.
1 Pet was adopted between 1 and 7 days (1st week). 
2 Pet was adopted between 8 and 30 days (1st month) after being listed.
3 Pet was adopted between 31 and 90 days (2nd & 3rd month) after being listed.
4 No adoption after 100 days of being listed.  
We transformed AdoptionSpeed to numerical variable: ASnum.
```{r ASnum}
ggplot(data=dat, aes(dat$ASnum)) + 
  geom_histogram(breaks=seq(0, 100, by=1),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="Histogram of ASnum", x="Pet adopted days") + 
  scale_fill_gradient("Count", low="green", high="red")

#Q-Q plot
qqnorm(dat$ASnum, main="Q-Q plot of ASnum of pets", ylab = "days",col="blue")
qqline(dat$ASnum)
```
This is Age.  
```{r Age}
tmpDatAge = dat
tmpDatAge$Age = as.numeric(dat$Age)

ggplot(data=tmpDatAge, aes(Age)) + 
  geom_histogram(breaks=seq(0, 100, by=5),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="Histogram of Age of pets", x="Age(in months)") + 
  scale_fill_gradient("Count", low="green", high="red")
#Boxplot
ggplot(data = tmpDatAge,aes(y = Age))+
  geom_boxplot() + 
  geom_boxplot( colour="orange", fill="#7777cc", outlier.colour="red", outlier.shape=8, outlier.size=4) +
  labs(title="Age boxplot using `ggplot`",x="", y = "Age")

#Q-Q plot
qqnorm(tmpDatAge$Age, main="Q-Q plot of Age of pets", ylab = "Age(in months)",col="blue")
qqline(tmpDatAge$Age)
```
This is photoAmt: number of photo  
```{r PhotoAmt}
tmpPhotoAmt = dat
tmpPhotoAmt$PhotoAmt = as.numeric(tmpPhotoAmt$PhotoAmt)

ggplot(data=tmpPhotoAmt, aes(PhotoAmt)) + 
  geom_histogram(breaks=seq(0, 30, by=2),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="Histogram of PhotoAmt of pets") + 
  scale_fill_gradient("Count", low="green", high="red")

#Boxplot
ggplot(data = tmpPhotoAmt,aes(y = PhotoAmt))+
  geom_boxplot() + 
  geom_boxplot( colour="orange", fill="#7777cc", outlier.colour="red", outlier.shape=8, outlier.size=4) +
  labs(title="PhotoAmt boxplot using `ggplot`",x="", y = "Number of Photoes")

#Q-Q plot
qqnorm(tmpPhotoAmt$PhotoAmt, main="Q-Q plot of PhotoAmt of pets", ylab = "PhotoAmt",col="blue")
qqline(tmpPhotoAmt$PhotoAmt)
```

### Categorical variable  
Let's us focusing on some categorical variables.  
The first one is Type.  
```{r Type}
dog = nrow(dat[dat$Type=="1",])
cat = nrow(dat[dat$Type=="2",])
totalamount = dog+cat

percentDog = format(dog/totalamount * 100, digits = 4)
percentCat = format(cat/totalamount * 100, digits = 4)

Prop <- c(dog,cat)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(5, "RdYlGn") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Dog: ",percentDog,"%"), paste("Cat: ",percentCat,"%")),
    main = "Pie plot of Type of pets", border="white", col=myPalette )
```
The second one is Gender.  
```{r Gender}
#pie plot
male = nrow(dat[dat$Gender=="1",])
female = nrow(dat[dat$Gender=="2",])
percentMale = format(male/(male+female) * 100, digits = 4)
percentFemale = format(female/(male+female) * 100, digits = 4)
Prop <- c(male,female)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(3, "Pastel1") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Male: ", percentMale,"%"),paste("Female: ",percentFemale,"%")), main = "Pie plot of Gender", border="white", col=myPalette )
```
The third one is MaturitySize.  
```{r MaturitySize}
small = nrow(dat[dat$MaturitySize=="1",])
medium = nrow(dat[dat$MaturitySize=="2",])
large = nrow(dat[dat$MaturitySize=="3",])
extraLarge = nrow(dat[dat$MaturitySize=="4",])
totalamount = small+medium+large+extraLarge

percentSmall = format(small/totalamount * 100, digits = 4)
percentMedium = format(medium/totalamount * 100, digits = 4)
percentLarge = format(large/totalamount * 100, digits = 4)
percentExtraLarge = format(extraLarge/totalamount * 100, digits = 4)
#pie plot
Prop <- c(small,medium,large,extraLarge)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(8, "Spectral") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Small: ",percentSmall,"%"), paste("Medium: ",percentMedium,"%"), paste("Large: ",percentLarge,"%"), paste("Extra Large: ",percentExtraLarge,"%")), main = "Pie plot of MaturitySize", border="white", col=myPalette )

```
The fourth one is FurLength.  
```{r FurLength}

short = nrow(dat[dat$FurLength=="1",])
medium = nrow(dat[dat$FurLength=="2",])
long = nrow(dat[dat$FurLength=="3",])
totalamount = short+medium+long

percentShort = format(short/totalamount * 100, digits = 4)
percentMedium = format(medium/totalamount * 100, digits = 4)
percentLong = format(long/totalamount * 100, digits = 4)


#pie plot
Prop <- c(short,medium,long)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(3, "Set2") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Short: ",percentShort,"%"), paste("Medium: ",percentMedium,"%"), paste("Long: ", percentLong,"%")), main = "Pie plot of FurLength", border="white", col=myPalette )
```
The fifth one is Vaccinated.  
```{r Vaccinated}
vaccinatedYes = nrow(dat[dat$Vaccinated=="1",])
vaccinatedNo = nrow(dat[dat$Vaccinated=="2",])
vaccinatedNotSure = nrow(dat[dat$Vaccinated=="3",])
totalamount = vaccinatedYes+vaccinatedNo+vaccinatedNotSure

percentVaccinatedYes = format(vaccinatedYes/totalamount * 100, digits = 4)
percentVaccinatedNo = format(vaccinatedNo/totalamount * 100, digits = 4)
percentVaccinatedNotSure = format(vaccinatedNotSure/totalamount * 100, digits = 4)

#pie plot
Prop <- c(vaccinatedYes,vaccinatedNo,vaccinatedNotSure)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(11,"Paired") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Vaccinated: ",percentVaccinatedYes, "%"), paste("Unvaccinated: ",percentVaccinatedNo, "%"), paste("Not Specified: ",percentVaccinatedNotSure, "%")), main = "Pie plot of Vaccinated", border="white", col=myPalette ) 
```


## Chapter 4

### SMART: Do dogs get adopted faster than cats?

First, we looked at animal type. In order to answer this question, we split the data into two groups, dogs and cats. Because there were only 2 samples we decided to use a two-sample t-test to evaluate if there was a significant difference between the average adoption speed of dogs and cats. 
```{r}
library('ggplot2')
library('lmtest')

# boxplot of the animal type adoption speed
ggplot(data = dat, aes(y = ASnum, x = Type, fill = Type)) + 
  geom_boxplot() +
  labs(title = "Adoption Speed by Animal Type", x = 'Animal Type', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Dog", "2" = "Cat")) + 
  scale_fill_discrete(labels = c("Dog", "Cat"))

```

```{r, results='markup'}

# running a variance check between dogs and cats
bp_dogscats = bptest(data = dat, formula = ASnum ~ Type)
bp_dogscats

# subsetting the data
dogs = subset(dat, dat$Type == 1) # only the dogs
cats = subset(dat, dat$Type == 2) # only the cats


# 2 sample t-test
dc_t = t.test(dogs$ASnum, cats$ASnum)
dc_t
```

The first thing we had to do was a check of the homogeneity of variance. To do this we did a Breusch-Pagan (BP) test. The null hypothesis for this test is that the two groups have the same variance, while the alternative hypothesis is that the groups have difference variances. Because the p_value is less than our alpha of 0.05 (p = `r bp_dogscats$p.value`) we have to reject the null hypothesis that the groups have the same variance. Because of this, the two-sample t-test is not technically a valid statistical test, however we went ahead and performed the test regardless. 

The p_value of the t-test was also less than our alpha of 0.05 (p = `r dc_t$p.value`), allowing us to reject our null hypothesis that there is no difference between the means of the two groups. Because of this we can conclude that there is a difference in adoption speed between cats and dogs. Because the mean adoption speed for dogs is higher (slower) than the mean adoption speed for cats, we can conclude that dogs get adopted slower than cats, and that animal type does have an effect on adoption speed.

### SMART: Do physical attributes affect adoption speed?

The next question we wanted to answer was what physical attributes affect adoption speed. To answer this question we looked at the sex of the animal, the size, the fur length, and the vaccination status. Begining with the sex of the animal, we compared male animals to female animals.

```{r, results='markup'}
### GENDER

# variance check for the gender
bp_mf = bptest(data = dat, formula = ASnum ~ Gender)
bp_mf

# subsetting the data for the t-test
male = subset(dat, dat$Gender == 1) # only the male dogs
female = subset(dat, dat$Gender == 2) # only the female dogs

# 2 sample t-test
mf_t = t.test(male$ASnum, female$ASnum)
mf_t

# box plot of the adoption speed by gender
ggplot(data = dat, aes(y = ASnum, x = Gender, fill = Gender)) + 
  geom_boxplot() + 
  labs(title = "Adoption Speed by Animal Gender", x = 'Animal Gender', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Male", "2" = "Female")) + 
  scale_fill_discrete(labels = c("Male", "Female"))
```

Again, the BP test of variance had a p_value that was significant for our alpha of 0.05 (p = `r bp_mf$p.value`), so we were able to reject the null hypothesis that the variance was equal between the groups. Again, this meant that a t-test was not appropriate, however we did a two-sample t-test regardless and found that this was also significant (p = `r mf_t$p.value`). So we can reject the null hypothesis that the mean adoption speed between male and female animals is the same. Because the mean adoption speed of male animals was less than the mean adoption speed than female animals, we can conclude that female animals get adopted quicker than male dogs.

```{r, results='markup'}
### SIZE

# variance check for the size
bp_size = bptest(data = dat, formula = ASnum ~ MaturitySize)
bp_size

# box plot of adoption speed and size
ggplot(data = dat, aes(y = ASnum, x = MaturitySize, fill = MaturitySize)) + 
  geom_boxplot() + 
  labs(title = "Adoption Speed by Maturity Size", x = 'Maturity Size', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Small", "2" = "Medium", "3" = "Large", "4" = "Extra Large")) + 
  scale_fill_discrete(labels = c("Small", "Medium", "Large", "Extra Large"))


# running an anova on the size
anova_size = anova(aov(data = dat, ASnum ~ MaturitySize))
anova_size


# running the post-hoc analysis
turkey_size = TukeyHSD(aov(data = dat, ASnum ~ MaturitySize))
turkey_size
```

Looking at size we have over 2 samples, so we chose to run an ANOVA to test if the mean adoption speed differed between different sized animals. Once again, we ran a BP test of variance that produced a significant p_value (p = `r bp_size$p.value`), so an ANOVA is technically not appropriate. However we did run the analysis regardless. We ended up with a significant p_value for the anova (p = `r anova_size$'Pr(>F)'`), so we ran a TukeyHSD post-hoc test and found that the means between the small and medium dogs, and the medium and large dogs, were significantly different. We can therefore conclude that animal size does affect adoption speed.

```{r, results='markup'}
### FURLENGTH

# variance check for fur length
bp_fur = bptest(data = dat, formula = ASnum ~ FurLength)
bp_fur

# fur length box plot
ggplot(data = dat, aes(y = ASnum, x = FurLength, fill = FurLength)) + 
  geom_boxplot() +
  labs(title = "Adoption Speed by Fur Length", x = 'Fur Length', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Short", "2" = "Medium", "3" = "Long")) + 
  scale_fill_discrete(labels = c("Short", "Medium", "Long"))

# anova for fur length
anova_fur = anova(aov(data = dat, ASnum ~ FurLength))
anova_fur

# post-hoc tests for fur length
turkey_fur = TukeyHSD(aov(data = dat, ASnum ~ FurLength))
turkey_fur
```

Looking next at fur length, once again our BP test of equal variance failed (p = `r bp_fur$p.value`). However we continued with our analysis. Because we have more than 2 samples we again decided to use an ANOVA to determine if the average adoption speed varied based on fur length. Our ANOVA did produce a significant result (p = `r anova_fur$'Pr(>F)'`) allowing us to reject the null hypothesis that the adoption speed was the same across fur lengths. Because we got a signficant result we again ran a TukeyHSD post-hoc analysis and found that the difference between each of the fur length groups was significant. We're able to conclude that fur length does impact adoption speed.

```{r, results='markup'}
### VACCINATED

# variance check for vaccination status
bp_vac = bptest(data = dat, formula = ASnum ~ Vaccinated)
bp_vac

# vaccination box plot
ggplot(data = dat, aes(y = ASnum, x = Vaccinated, fill = Vaccinated)) + 
  geom_boxplot() + 
  labs(title = "Adoption Speed by Vaccination Status", x = 'Vaccination Status', y = 'Adoption Speed') + 
  scale_x_discrete(labels=c("1" = "Vaccinated", "2" = "Unvaccinated", "3" = "Not Specified")) + 
  scale_fill_discrete(labels = c("Vaccinated", "Unvaccinated", "Not Specified"))

# vaccination anova
anova_vac = anova(aov(data = dat, ASnum ~ Vaccinated))
anova_vac

# post-hoc analysis for vaccination
turkey_vac = TukeyHSD(aov(data = dat, ASnum ~ Vaccinated))
turkey_vac
```

Finally, we looked at vaccination status. Once again we were able to reject the null hypothesis of the BP variance test (p = `r bp_vac$p.value`) but we continued with our analysis. We ran an ANOVA since there were 3 sample levels and again found a significant result (p = `r anova_vac$'Pr(>F)'`). In our post-hoc analysis we found that the difference between the vaccinated and unvaccinated, and the unvaccinated and unspecified group was significantly different. We can conclude that vaccination status did impact the adoption speed of the animal. 


The final conclusion from our analysis of categorical physical characteristics is that the difference in variance across the groups does not allow us to draw any meaningful conclusions. However, if we ignore that assumption for this sample of data, we are able to conclude that all the categorical physical characteristics, including the type of the animal, does impact the adoption speed of the animal.

# Chapter 5

### SMART: What numerical variables influence adoption speed?

We will run single variable OLS regressions for the three independent numerical variables.
First, age.

```{r results='markup'}
lm_age <- lm(ASnum ~ Age, data=dat )
table_lm_age <- tbl_regression(lm_age, intercept=TRUE)
table_lm_age
```
Clearly, age is statistically significant, as the p-value is below 0.05.

Lets look at VideoAmt next.

```{r results='markup'}
lm_video <- lm(ASnum ~ VideoAmt, data=dat)
table_lm_video <- tbl_regression(lm_video, intercept=TRUE)
table_lm_video
```

Finally, lets check PhotoAmt.

```{r results='markup'}
lm_photo <- lm(ASnum ~ PhotoAmt, data=dat)
table_lm_photo <-tbl_regression(lm_photo, intercept=TRUE)
table_lm_photo
```

As we can see, only Age and PhotoAmt results in a p-value less than 0.05.

Answer to SMART question 3: Age and PhotoAmt are the only two numerical variables which alone can be considered statistically significant. 

### SMART: What variables, both categorical and numerical, result in the best predictave model?

### Using feature selection tools to identify best model.

Starting with generating an OLS model including all the variables we are considering.
```{r results='markup'}
loadPkg("leaps")
reg_model <- regsubsets(ASnum~ Age + Gender + MaturitySize + FurLength + Vaccinated + VideoAmt + PhotoAmt, data = dat, nbest = 2, method = 'exhaustive')
```


Let us now create plots for all 4 of the methods we want to use. Exhaustive, forward, backward, and sequential methods. 
```{r results='markup'}
plot(reg_model, scale = "adjr2", main = "Adjusted R^2 Exhaustive")
reg_model_forward <- regsubsets(ASnum~ Age + Gender + MaturitySize + FurLength + Vaccinated + VideoAmt + PhotoAmt, data = dat, nbest = 2, method = 'forward')
plot(reg_model_forward, scale = "adjr2", main = "Adjusted R^2 Forward")
reg_model_backward <- regsubsets(ASnum~ Age + Gender + MaturitySize + FurLength + Vaccinated + VideoAmt + PhotoAmt, data = dat, nbest = 2, method = 'backward')
plot(reg_model_backward, scale = "adjr2", main = "Adjusted R^2 Backward")
reg_model_sequential <- regsubsets(ASnum~ Age + Gender + MaturitySize + FurLength + Vaccinated + VideoAmt + PhotoAmt, data = dat, nbest = 2, method = 'seqrep')
plot(reg_model_sequential, scale = "adjr2", main = "Adjusted R^2 Sequential Replacement")
```
No difference in variables between exhaustive, forward, and backward selection:
Age, Gender 2, MaturitySize2, MaturitySize3, FurLength2, Furlength3, Vaccinated2, and PhotoAmt

Sequential Replacement recommended:
Gender2, MaturitySize2, FurLength2, FurLength3, Vaccinated2, PhotoAmt

Dropping Age, MaturitySize3 between two models


```{r results='markup'}
dat_drop_mat_4 <- subset(dat, ! dat$MaturitySize == 4)
dat_drop_mat_4 <- subset(dat_drop_mat_4, ! dat_drop_mat_4$Vaccinated== 3 )
dat_drop_mat_4_3 <- subset(dat_drop_mat_4, ! dat_drop_mat_4$MaturitySize == 3)
exh_for_back_model <- lm(ASnum~ Age + Gender + MaturitySize + FurLength + Vaccinated + PhotoAmt, data=dat_drop_mat_4)
table_exh_for_back_model <- tbl_regression(exh_for_back_model, intercept=TRUE)
add_vif(table_exh_for_back_model, statistic = "GVIF")
```

```{r results='markup'}
sequential_model <- lm(ASnum ~ Gender + MaturitySize + FurLength + Vaccinated + PhotoAmt, data=dat_drop_mat_4_3)
table_sequential_model <- tbl_regression(sequential_model, intercept=TRUE)
add_vif(table_sequential_model, statistic = "GVIF")
```