---
title: "Independent_Variables_EDA_Final"
author: "YixiLiang"
date: "2021/10/31"
output:
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# The package "ezids" (EZ Intro to Data Science) includes a lot of the helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
library(ezids)
 
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```


## Intro
```{r}
## Loading data
data = read.csv('datafile.csv')
data$AdoptionSpeed <- factor(data$AdoptionSpeed, order=T, levels = c(0,1,2,3,4))
# converting the data types we want
data$Type = as.factor(data$Type)
data$MaturitySize = as.factor(data$MaturitySize)
data$FurLength = as.factor(data$FurLength)
data$Vaccinated = as.factor(data$Vaccinated)
data$Gender = as.factor(data$Gender)
# Only looking at profiles with 1 animal
data = subset(data, data$Quantity == 1)
data = subset(data, ! data$AdoptionSpeed == 4)
# Only pulling the columns we want to look at
dat = data[c('Type', 'Age', 'Gender', 'MaturitySize', 'FurLength', 'Vaccinated', 'PhotoAmt', 'AdoptionSpeed')]
#summary(dat)
# Adding random values
dat$ASnum = NaN
dat$ASnum[dat$AdoptionSpeed == 0] = 0
dat$ASnum[dat$AdoptionSpeed == 1] = runif(sum(dat$AdoptionSpeed == 1), 1, 8)
dat$ASnum[dat$AdoptionSpeed == 2] = runif(sum(dat$AdoptionSpeed == 2), 8, 31)
dat$ASnum[dat$AdoptionSpeed == 3] = runif(sum(dat$AdoptionSpeed == 3), 31, 91)
#dat$ASnum[dat$AdoptionSpeed == 4] = 100
#summary(dat)
```

```{r}
loadPkg("ggplot2")
```
## Chapter 3
### Numerical variable
In this chapter we will introduce some variables. The first one is AdoptionSpeed.  
```{r histogram of AdoptionSpeed}
loadPkg("ggplot2")
#Histogram
ggplot(data=dat, aes(AdoptionSpeed)) + 
  geom_bar(breaks=seq(0, 4, by=1),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="AdoptionSpeed of pets", x = "AdoptionSpeed") + 
  scale_fill_gradient("Count", low="green", high="red")

```
AdoptionSpeed is divided into 4 levels.   
0 means pet was adopted on the same day as it was listed.
1 Pet was adopted between 1 and 7 days (1st week). 
2 Pet was adopted between 8 and 30 days (1st month) after being listed.
3 Pet was adopted between 31 and 90 days (2nd & 3rd month) after being listed.
4 No adoption after 100 days of being listed.  
We transformed AdoptionSpeed to numerical variable: ASnum.
```{r ASnum}
ggplot(data=dat, aes(dat$ASnum)) + 
  geom_histogram(breaks=seq(0, 100, by=1),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="Histogram of ASnum", x="Pet adopted days") + 
  scale_fill_gradient("Count", low="green", high="red")

#Q-Q plot
qqnorm(dat$ASnum, main="Q-Q plot of ASnum of pets", ylab = "days",col="blue")
qqline(dat$ASnum)
```
This is Age.  
```{r Age}
tmpDatAge = dat
tmpDatAge$Age = as.numeric(dat$Age)

ggplot(data=tmpDatAge, aes(Age)) + 
  geom_histogram(breaks=seq(0, 100, by=5),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="Histogram of Age of pets", x="Age(in months)") + 
  scale_fill_gradient("Count", low="green", high="red")
#Boxplot
ggplot(data = tmpDatAge,aes(y = Age))+
  geom_boxplot() + 
  geom_boxplot( colour="orange", fill="#7777cc", outlier.colour="red", outlier.shape=8, outlier.size=4) +
  labs(title="Age boxplot using `ggplot`",x="", y = "Age")

#Q-Q plot
qqnorm(tmpDatAge$Age, main="Q-Q plot of Age of pets", ylab = "Age(in months)",col="blue")
qqline(tmpDatAge$Age)
```
This is photoAmt: number of photo  
```{r PhotoAmt}
tmpPhotoAmt = dat
tmpPhotoAmt$PhotoAmt = as.numeric(tmpPhotoAmt$PhotoAmt)

ggplot(data=tmpPhotoAmt, aes(PhotoAmt)) + 
  geom_histogram(breaks=seq(0, 30, by=2),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="Histogram of PhotoAmt of pets") + 
  scale_fill_gradient("Count", low="green", high="red")

#Boxplot
ggplot(data = tmpPhotoAmt,aes(y = PhotoAmt))+
  geom_boxplot() + 
  geom_boxplot( colour="orange", fill="#7777cc", outlier.colour="red", outlier.shape=8, outlier.size=4) +
  labs(title="PhotoAmt boxplot using `ggplot`",x="", y = "Number of Photoes")

#Q-Q plot
qqnorm(tmpPhotoAmt$PhotoAmt, main="Q-Q plot of PhotoAmt of pets", ylab = "PhotoAmt",col="blue")
qqline(tmpPhotoAmt$PhotoAmt)
```

### Categorical variable  
Let's us focusing on some categorical variables.  
The first one is Type.  
```{r Type}
dog = nrow(dat[dat$Type=="1",])
cat = nrow(dat[dat$Type=="2",])
totalamount = dog+cat

percentDog = format(dog/totalamount * 100, digits = 4)
percentCat = format(cat/totalamount * 100, digits = 4)

Prop <- c(dog,cat)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(5, "RdYlGn") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Dog: ",percentDog,"%"), paste("Cat: ",percentCat,"%")),
    main = "Pie plot of Type of pets", border="white", col=myPalette )
```
The second one is Gender.  
```{r Gender}
#pie plot
male = nrow(dat[dat$Gender=="1",])
female = nrow(dat[dat$Gender=="2",])
percentMale = format(male/(male+female) * 100, digits = 4)
percentFemale = format(female/(male+female) * 100, digits = 4)
Prop <- c(male,female)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(3, "Pastel1") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Male: ", percentMale,"%"),paste("Female: ",percentFemale,"%")), main = "Pie plot of Gender", border="white", col=myPalette )
```
The third one is MaturitySize.  
```{r MaturitySize}
small = nrow(dat[dat$MaturitySize=="1",])
medium = nrow(dat[dat$MaturitySize=="2",])
large = nrow(dat[dat$MaturitySize=="3",])
extraLarge = nrow(dat[dat$MaturitySize=="4",])
totalamount = small+medium+large+extraLarge

percentSmall = format(small/totalamount * 100, digits = 4)
percentMedium = format(medium/totalamount * 100, digits = 4)
percentLarge = format(large/totalamount * 100, digits = 4)
percentExtraLarge = format(extraLarge/totalamount * 100, digits = 4)
#pie plot
Prop <- c(small,medium,large,extraLarge)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(8, "Spectral") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Small: ",percentSmall,"%"), paste("Medium: ",percentMedium,"%"), paste("Large: ",percentLarge,"%"), paste("Extra Large: ",percentExtraLarge,"%")), main = "Pie plot of MaturitySize", border="white", col=myPalette )

```
The fourth one is FurLength.  
```{r FurLength}

short = nrow(dat[dat$FurLength=="1",])
medium = nrow(dat[dat$FurLength=="2",])
long = nrow(dat[dat$FurLength=="3",])
totalamount = short+medium+long

percentShort = format(short/totalamount * 100, digits = 4)
percentMedium = format(medium/totalamount * 100, digits = 4)
percentLong = format(long/totalamount * 100, digits = 4)


#pie plot
Prop <- c(short,medium,long)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(3, "Set2") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Short: ",percentShort,"%"), paste("Medium: ",percentMedium,"%"), paste("Long: ", percentLong,"%")), main = "Pie plot of FurLength", border="white", col=myPalette )
```
The fifth one is Vaccinated.  
```{r Vaccinated}
vaccinatedYes = nrow(dat[dat$Vaccinated=="1",])
vaccinatedNo = nrow(dat[dat$Vaccinated=="2",])
vaccinatedNotSure = nrow(dat[dat$Vaccinated=="3",])
totalamount = vaccinatedYes+vaccinatedNo+vaccinatedNotSure

percentVaccinatedYes = format(vaccinatedYes/totalamount * 100, digits = 4)
percentVaccinatedNo = format(vaccinatedNo/totalamount * 100, digits = 4)
percentVaccinatedNotSure = format(vaccinatedNotSure/totalamount * 100, digits = 4)

#pie plot
Prop <- c(vaccinatedYes,vaccinatedNo,vaccinatedNotSure)
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(11,"Paired") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c(paste("Vaccinated: ",percentVaccinatedYes, "%"), paste("Unvaccinated: ",percentVaccinatedNo, "%"), paste("Not Specified: ",percentVaccinatedNotSure, "%")), main = "Pie plot of Vaccinated", border="white", col=myPalette )
```
