---
title: "Independent_Variables_EDA"
author: "YixiLiang"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# The package "ezids" (EZ Intro to Data Science) includes a lot of the helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
library(ezids)
 
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```


## Loading data
```{r}
#dat = read.csv("E:/Users/REX/RstudioProjects/data_torture/chapter_3/pets_test.csv")
dat <- data.frame(read.csv('datafile.csv'))
dat$AdoptionSpeed <- factor(dat$AdoptionSpeed, order=T, levels = c(0,1,2,3,4))
#data = na.omit(dat) #delete na from data set


dat$Type = as.factor(dat$Type)
dat$Name = as.factor(dat$Name)
dat$Breed1 = as.factor(dat$Breed1)
dat$Breed2 = as.factor(dat$Breed2)
dat$Gender = as.factor(dat$Gender)
dat$Color1 = as.factor(dat$Color1)
dat$Color2 = as.factor(dat$Color2)
dat$Color3 = as.factor(dat$Color3)
dat$MaturitySize = as.factor(dat$MaturitySize)
dat$FurLength = as.factor(dat$FurLength)
dat$Vaccinated = as.factor(dat$Vaccinated)
dat$Dewormed = as.factor(dat$Dewormed)
dat$Sterilized = as.factor(dat$Sterilized)
dat$Health = as.factor(dat$Health)
dat$State = as.factor(dat$State)

summary(dat)
```
```{r}
dat = subset(dat, dat$Quantity == 1)
dat = subset(dat, ! dat$AdoptionSpeed == 4)
summary(dat)
```

```{r}
dat$ASnum = NaN
dat$ASnum[dat$AdoptionSpeed == 0] = 0
<<<<<<< Updated upstream
# dat$ASnum[dat$AdoptionSpeed == 1] = runif(sum(dat$AdoptionSpeed == 1), 1, 7)
# dat$ASnum[dat$AdoptionSpeed == 2] = runif(sum(dat$AdoptionSpeed == 2), 8, 30)
# dat$ASnum[dat$AdoptionSpeed == 3] = runif(sum(dat$AdoptionSpeed == 3), 31, 90)
#dat$ASnum[dat$AdoptionSpeed == 4] = runif(sum(dat$AdoptionSpeed == 4), 100, 200)
#rchisq(sum(dat$AdoptionSpeed == 4), ?, ncp = 0)

#Maybe this one is better
dat$ASnum[dat$AdoptionSpeed == 1] = 1 + rgeom(sum(dat$AdoptionSpeed == 1), 0.02)
dat$ASnum[dat$AdoptionSpeed == 2] = 8 + rgeom(sum(dat$AdoptionSpeed == 2), 0.02)
dat$ASnum[dat$AdoptionSpeed == 3] = 31 + rgeom(sum(dat$AdoptionSpeed == 3), 0.02)
#dat$ASnum[dat$AdoptionSpeed == 4] = 100 + rgeom(sum(dat$AdoptionSpeed == 4), 0.02)
summary(dat)
```
```{r}
#The second method 
dat$ASnum[dat$AdoptionSpeed == 1] = 1 * rgeom(sum(dat$AdoptionSpeed == 1), 0.1)
dat$ASnum[dat$AdoptionSpeed == 2] = 2 * rgeom(sum(dat$AdoptionSpeed == 2), 0.1)
dat$ASnum[dat$AdoptionSpeed == 3] = 3 * rgeom(sum(dat$AdoptionSpeed == 3), 0.1)
#dat$ASnum[dat$AdoptionSpeed == 4] = 4 * rgeom(sum(dat$AdoptionSpeed == 4), 0.1)

#This function rgeom(n, prob) is Geometric Distribution, random generation for the geometric distribution with parameter prob 
summary(dat)
```
=======

dat$ASnum[dat$AdoptionSpeed == 1] = runif(sum(dat$AdoptionSpeed == 1), 1, 7)
dat$ASnum[dat$AdoptionSpeed == 2] = runif(sum(dat$AdoptionSpeed == 2), 8, 30)
dat$ASnum[dat$AdoptionSpeed == 3] = runif(sum(dat$AdoptionSpeed == 3), 31, 90)
#summary(dat)
```

>>>>>>> Stashed changes

```{r histogram of AdoptionSpeed}
loadPkg("ggplot2")
tmpAdoptionSpeed = as.numeric(dat$AdoptionSpeed)
ggplot(data=dat, aes(tmpAdoptionSpeed)) + 
  geom_histogram(breaks=seq(0, 4, by=1),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="AdoptionSpeed of pets") + 
  scale_fill_gradient("Count", low="green", high="red")

#Q-Q plot
qqnorm(tmpAdoptionSpeed, main="Q-Q plot of AdoptionSpeed of pets", ylab = "days",col="blue")
qqline(tmpAdoptionSpeed)
```

```{r histogram of ASnum}
loadPkg("ggplot2")
ggplot(data=dat, aes(dat$ASnum)) + 
  geom_histogram(breaks=seq(0, 200, by=5),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="ASnum of pets") + 
  scale_fill_gradient("Count", low="green", high="red")

#Q-Q plot
qqnorm(dat$ASnum, main="Q-Q plot of ASnum of pets", ylab = "days",col="blue")
qqline(dat$ASnum)

```

```{r histogram of Fee}
loadPkg("ggplot2")
tmpFee = as.numeric(dat$Fee)
ggplot(data=dat, aes(tmpFee)) + 
  geom_histogram(breaks=seq(0, 500, by=50),
                 col="red",
                 aes(fill=..count..)) +
  labs(title="Feed of pets") + 
  scale_fill_gradient("Count", low="green", high="red")

#Q-Q plot
qqnorm(tmpFee, main="Q-Q plot of ASnum of pets", ylab = "days",col="blue")
qqline(tmpFee)
```

